---
title: "PSV Analysis"
author: DIME and PSV Safety Awards
output: 
  html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---

This document summarizes results from piloting the Matatu Rider Feedback project. So far, there have been 2 pilots:

* __Pilot 1__ 
  * 15 Vehicles involved. 
  * 8 vehicles had stickers advertising that riders receive 200 KES after completing the survey 
  * 7 vehicles advertised receiving 100 KES after completing the survey.
  * We received a large number of responses after only a couple days of running the pilot. Consequently, we lowered the award for types to 20 KES. The lower award was advertised in the survey. 
  * In the shortcode, to identify the vehicle, users entered a 2 digit number that was put on the stickers. 
  * The pilot started on June 28, and the stickers indicated that the award expired on July 20.

* __Pilot 2__ 
  * 15 vehicles involved. 
  * 9 vehicles had stickers advertising that riders received 50 KES after completing the survey (6 had both the shortcode and QR code, while 3 only had the QR code)  
  * 6 vehicles advertised winning 50 KES in airtime. 
  * Instead of a two digit number, the licence plate was put on the sticker and riders entered the plate number in the shortcode survey.
  * As with the previous pilot, we received a large number of responses in the first couple days. We paused the shortcode survey where a rider automatically recieved the 50 KES award due to large costs. 
  * On August 12, we restarted the survey -- only asking 3 questions and lowered the award to 10 KES.
  * The stickers were installedon July 28 and the stickers indicate that the award offer expires on August 31.


```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r dataprep, echo=FALSE, include=FALSE}
# Setup ------------------------------------------------------------------------
if(Sys.info()[["user"]] == "robmarty") github <- "~/Documents/Github/PSV-Rider-Feedback/"

source(file.path(github, "_psv_feedback_master.R"))

# Parameters -------------------------------------------------------------------
PILOT_1_DATE <- "2020-06-26"
PILOT_2_DATE <- "2020-07-28"

# Load Data --------------------------------------------------------------------
data <- readRDS(file.path(dropbox_file_path, "Data", "Rider Feedback",
                          "Echo Mobile Data", "FinalData", "echo_data.Rds"))

# Prep Data --------------------------------------------------------------------
data <- data %>%
  filter(date >= "2020-06-26")

```

# Number of Responses

## Daily - By Award Type
```{r, fig.height=4}

data_award <- data %>%
  filter(!is.na(complete_date)) %>%
  group_by(date, award_actual_getwin) %>%
  dplyr::summarise(N = n()) 

data_total <- data %>%
  filter(!is.na(complete_date)) %>%
  group_by(date) %>%
  dplyr::summarise(N_total = n()) %>%
  ungroup() 

ggplot() +
  geom_col(data = data_award, 
           aes(x=date, y=N,
               fill = award_actual_getwin,
               group = award_actual_getwin)) +
  geom_text(data = data_total,
            aes(x=date, y=N_total,
                label = N_total), nudge_y = 50, color = "black", size=2) +
  geom_vline(xintercept = as.Date(PILOT_2_DATE),
             color = "red") +
  labs(fill = "Award") +
  theme_ipsum()
```


## Overall - Award Type

TODO: Standardize by days award active.

```{r}

data %>% 
  group_by(award_actual_getwin) %>%
  summarise(N = n()) %>%
  ggplot(aes(x = award_actual_getwin, y = N)) + 
  geom_col(alpha = 0.9, fill = "aquamarine") +
  geom_text(aes(label = N)) +
  labs(x = "Award Type") +
  coord_flip() +
  theme_ipsum()

```



## Daily - By Vehicle
```{r}

data_veh <- data %>%
  filter(!is.na(complete_date)) %>%
  group_by(date, matatu_no_clean) %>%
  dplyr::summarise(N = n()) 

data_total <- data %>%
  filter(!is.na(complete_date)) %>%
  group_by(date) %>%
  dplyr::summarise(N_total = n()) %>%
  ungroup() 

ggplot() +
  geom_col(data = data_veh, 
           aes(x=date, y=N,
               fill = matatu_no_clean,
               group = matatu_no_clean)) +
  geom_text(data = data_total,
            aes(x=date, y=N_total,
                label = N_total), nudge_y = 50, color = "black", size=2) +
  geom_vline(xintercept = as.Date(PILOT_2_DATE),
             color = "red") +
  labs(fill = "Award") +
  theme_ipsum()

```

## Vehicle - Overall
```{r}

data %>%
  filter(!is.na(complete_date)) %>%
  filter(!is.na(matatu_no)) %>%
  group_by(matatu_no_clean) %>%
  dplyr::summarise(N = n()) %>%
  ggplot(aes(x = reorder(matatu_no_clean, N), y = N)) +
  geom_col(fill = "deepskyblue", alpha=0.8) +
  geom_text(aes(label = N), nudge_y = 50) +
  labs(x = "Matatu", y = "N") +
  coord_flip() +
  theme_ipsum()

```

## Hourly - By Vehicle
```{r, fig.height=20}
data %>%
    filter(!is.na(complete_date)) %>%
  mutate(datetime = complete_date %>% round_date(unit = "hour")) %>%
  filter(!is.na(datetime)) %>%
  group_by(datetime, week_date, matatu_no_clean) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  mutate(week_date = paste("Week of", week_date)) %>%
  ggplot() +
  geom_col(aes(x = datetime, y = N,
               group = matatu_no_clean,
               fill = matatu_no_clean)) +
  labs(title = "Completed Responses per Hour",
       fill = "Award") +
  facet_wrap(~week_date, ncol = 1, scale = "free_x") +
  theme_minimal()
```


## Response by Hour Histogram
```{r, fig.height=20}

data %>%
    filter(!is.na(complete_date)) %>%
  group_by(hour, week_date, matatu_no_clean) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  mutate(week_date = paste("Week of", week_date)) %>%
  ggplot() +
  geom_col(aes(x = hour, y = N,
               fill = matatu_no_clean,
               group = matatu_no_clean)) +
  labs(title = "Completed Responses per Hour") +
  facet_wrap(~week_date, ncol = 1, scale = "free_x") +
  theme_minimal()
```


## When enter reg no, how incorrect?
```{r}

data %>%
  filter(how_identif %in% "reg no",
         !is.na(reg_no_raw)) %>%
  group_by(reg_no_closest_dist) %>%
  dplyr::summarise(N = n()) %>%
  mutate(percent = (N/sum(N)*100) %>% round(1) %>% paste("%")) %>%
  ggplot(aes(x = reg_no_closest_dist, y = N)) +
  geom_col(fill = "dodgerblue", color = "black") +
  geom_text(aes(label = percent), nudge_y = 100) +
  labs(x = "Character difference to valid plate number",
       title = "Number of character differences to valid plate number") +
  theme_ipsum()


```

# Questions by Vehicle
```{r, fig.height=20}

data_long <- data %>%
  dplyr::select(matatu_no_clean, 
                driver_rating, speed_rating, occupancy, covid_measures) %>%
  pivot_longer(-matatu_no_clean) %>%
  
  filter(value != "") %>%
  
  # Number of responses per question, per vehicle
  group_by(matatu_no_clean, name) %>%
  mutate(N_veh = n()) %>%
  ungroup() %>%
  
  # Aggregate
  group_by(name, value, matatu_no_clean, N_veh) %>%
  summarise(N = n()) %>%
  ungroup()  %>%
  
  # Clean variables
  mutate(matatu_no_clean = paste0(matatu_no_clean, " - ", N_veh)) %>%
  mutate(N = N/N_veh)

data_long$name[data_long$name %in% "driver_rating"] <- "How would you rate your Matatu driver?"
data_long$name[data_long$name %in% "covid_measures"] <- "Were measures taken to prevent spread of COVID-19?"
data_long$name[data_long$name %in% "occupancy"] <- "On the Matatu, are there"
data_long$name[data_long$name %in% "speed_rating"] <- "How would you describe your Matatu driver's speed?"

data_long$value <- data_long$value %>% str_squish()

data_long$value <- factor(data_long$value, levels = c("No",
                                                      "Yes, but seemed limited",
                                                      "Yes, effective",
                                                      
                                                      "Very Unsafe",
                                                      "Unsafe",
                                                      "Safe",
                                                      "Very Safe",
                                                      
                                                      "Less people than seats",
                                                      "Same number of people as seats",
                                                      "More people than seats",
                                                      "More people than can fit",
                                                      
                                                      "Too slow",
                                                      "Okay",
                                                      "Fast",
                                                      "Dangerously Fast"))

out_plot <- lapply(data_long$name %>% unique, function(i){
  
  data_long %>%
    filter(name %in% i) %>%
    ggplot(aes(x = reorder(matatu_no_clean, N_veh), # 
               y = N, group=value, fill=value)) +
    geom_col(color="black") +
    labs(x="", y="",
         title = i) +
    theme_minimal() +
    coord_flip() +
    labs(x = "Matatu & Number of Responses", y = "Proportion", fill = "") 
  
}) 

do.call(grid.arrange, c(out_plot, ncol=1))
```

# Feedback

## Number of Words - Histogram
```{r}
# TO DO: Put into buckets and show % -- use geom_col

data$feedback_nwords_bin <- NA
data$feedback_nwords_bin[data$feedback_nwords %in% 1] <- "1"
data$feedback_nwords_bin[data$feedback_nwords %in% 2] <- "2"
data$feedback_nwords_bin[data$feedback_nwords %in% 3:5] <- "3 - 5"
data$feedback_nwords_bin[data$feedback_nwords %in% 6:10] <- "6 - 10"
data$feedback_nwords_bin[data$feedback_nwords %in% 11:20] <- "11 - 20"
data$feedback_nwords_bin[data$feedback_nwords %in% 21:9999] <- ">20"

data$feedback_nwords_bin <- data$feedback_nwords_bin %>%
  factor(levels = c("1", "2", "3 - 5", 
                    "6 - 10", "11 - 20", ">20"))

data %>%
  filter(feedback_nwords > 0) %>%
  group_by(feedback_nwords_bin) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  mutate(percent = (N/sum(N)*100) %>% round(1) %>% paste("%")) %>%
  ggplot(aes(x = feedback_nwords_bin,
             y = N)) +
  geom_col(fill = "dodgerblue", color = "black") +
  geom_text(aes(label = percent), nudge_y = 100) +
  labs(x = "N Words", 
       y = "N Responses") +
  theme_ipsum()
```

## Responses - All

```{r}
data %>%
  filter(feedback_nwords > 0) %>%
  arrange(desc(feedback_nwords)) %>%
  dplyr::select(feedback, matatu_no_clean) %>%
  dplyr::rename("Matatu" = matatu_no_clean,
                Feedback = feedback) %>%
  paged_table()
```

## Responses - Driver Unsafe or Very Unsafe

```{r}
data %>%
  filter(feedback_nwords > 0,
         driver_rating %in% c("Unsafe", "Very Unsafe")) %>%
  arrange(desc(feedback_nwords)) %>%
  dplyr::select(feedback, matatu_no_clean) %>%
  dplyr::rename("Matatu" = matatu_no_clean,
                Feedback = feedback) %>%
  paged_table()
```



# Number of Questions Answered

## Overall - 6 Qs Survey

```{r}
data %>%
  filter(N_questions %in% 6) %>%
  group_by(N_qs_answered) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  dplyr::mutate(prop = (round(N / sum(N),2)*100) %>% paste0("%")) %>%
  ggplot(aes(x = N_qs_answered, y = N)) +
  geom_col(fill = "deepskyblue4", color = "black") +
  geom_text(aes(label = prop), nudge_y = 150) +
  labs(x = "Number of Questions Answered") +
  theme_ipsum()

```

## By Vehicle - 6 Qs Survey

```{r}

data %>%
  
  filter(N_questions %in% 6) %>%
  
  mutate(N_qs_answered = N_qs_answered %>% as.factor() %>% fct_rev) %>%
  
  # Number of responses per question, per vehicle
  group_by(matatu_no_clean) %>%
  mutate(N_veh = n()) %>%
  ungroup() %>%
  
  group_by(N_qs_answered, matatu_no_clean, N_veh) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  
  mutate(prop = N / N_veh) %>%
  
  mutate(matatu_no_clean = paste0(matatu_no_clean, " - ", N_veh)) %>%
  
  ggplot(aes(x = reorder(matatu_no_clean, N_veh), y = prop, 
             group = N_qs_answered,
             fill = N_qs_answered)) +
  geom_col(color = "black") +
  labs(x = "Vehicle & Number of Responses", y = "Proportion",
       fill = "Number of\nQuestions Answered") +
  coord_flip() +
  theme_ipsum()
```

## Overall - 3 Qs Survey

After realizing we are low on funds after the 50 KES, we lowered the award to 10 KES and the number of questions to 3. This is clearly indicated at the start of the survey. The first message received in the shortcode is:

_"Welcome to this 3 question survey. \*\*We've lowered the award to 10 KES\*\*_

_What is the matatu reg. number? (Written on sticker)."_

```{r}
data %>%
  filter(N_questions %in% 3) %>%
  group_by(N_qs_answered) %>%
  dplyr::summarise(N = n()) %>%
  ungroup() %>%
  dplyr::mutate(prop = (round(N / sum(N),2)*100) %>% paste0("%")) %>%
  ggplot(aes(x = N_qs_answered, y = N)) +
  geom_col(fill = "deepskyblue4", color = "black") +
  geom_text(aes(label = prop), nudge_y = 1) +
  labs(x = "Number of Questions Answered") +
  theme_ipsum()

```






