---
title: "Sensor data description"
author: "DIME"
date: "13/12/2020"
output: 
 html_document:
    toc: true
    toc_depth: 3
    toc_float: true
    code_folding: hide
    toc_fold: TRUE
    highlight: tango
    keep_md: yes
    theme: cosmo
    number_sections: true
---

# Summary

* This document describes the sensor data for the matatu the sensor is placed in. We have been testing two sensors since September. Over the last 4 weeks, we have been trying to test different platforms to download the data from in order to get a dataset at the vehicle-time level which updates every 10 seconds. \
* Since we have been moving platforms, here we show the summary of 10 days of data in the last week of October and 2nd week for November.

```{r dataprep, echo=FALSE, include=FALSE}

#install.packages("shiny")
#install.packages("shinydashboard")
#install.packages("doBy")
#install.packages("kableExtra")

# Load all packages
library(dplyr)
library(leaflet)
library(lubridate)
library(purrr)
library(shiny)
library(shinydashboard)
library(raster)
library(tidyverse)
library(dplyr)
library(lubridate)
library(gridExtra)
library(hrbrthemes)
library(readxl)
library(labelled)
library(haven)
library(stringr)
library(rmarkdown)
library(wesanderson)
library(ggplot2)
library(dplyr)
library(doBy)
library(geojsonsf)
library(sf)
library(knitr)
library(kableExtra)
library(ggmap)
library(RColorBrewer)


# File path
if(Sys.info()[["user"]] == "meyhar") dropbox_file_path <- "~/Dropbox/PSV Rider Feedback"

#Sensor data
sensor_data <- file.path(dropbox_file_path, "Data", "Sensors", "pilot", "RawData")
output_sensor <- file.path(dropbox_file_path, "Data", "Sensors", "pilot", "Outputs")

# Load sensor data

sensor <- read_excel(file.path(sensor_data,"Matt_KBA_970C_Lions-_Trip_and_Eco-driving_detalization_19.11.2020_23-45-06.xlsx"), sheet = 5 )

sensor <- sensor %>%
  separate(col =Coordinates,
           into = c("lat", "lon"),
           sep = ",") %>%
  mutate(lat = lat %>% as.numeric(),
         lon = lon %>% as.numeric()) %>%
  mutate(Time = Time %>% dmy_hms()) %>%
  mutate(date = Time %>% as.Date())

sensor$hour <- lubridate::hour(sensor$Time)

```

# Number of days of data
* Data for days:\
* -- Oct 22- Oct 27 and,\
  -- Nov 11,12,13,19.\
* The sensor is on Alfred's car and a Matatu. This document has the information on the matatu.
* Gaps in data collection exists since we were trying 2-3 different platforms to store data on.
* Overall, the main variables of interest for us are the: \
    -lat/lon, \
    -speed,  \
    -harsh braking (dummy and magnitude) \
* Lat/lon, speed seem to be alright and making sense overall. 
* We need to bring up the following:
  -- Dates for which there are 0 values for speed and other variables, \
  -- Meaning if harsh braking/harsh acceleration is 0 \

``````{r, fig.height=8}
# Speed over time
sensor_data_days <- ggplot(sensor, aes(x=Time, y=Speed)) +
  geom_line(color="#69b3a2",  linetype=2) +
theme_ipsum() 
```
# Variable summary

## Speed
* We suspect the 0 in speed can be a combination of :\
  -- Less matatu movement due to covid, \
  -- They were having some technical issue in the device in the last week of October, first week of November due to which the device was not collecting any data. \
  -- We will make sure to check the reasons for this with Sam.
  
```{r}
sensor$speed_dum <- ifelse(sensor$Speed == 0, 1, 0)
kable(table(sensor$date, sensor$speed_dum), col.names = c("Non-zero speed", "Zero speed"), caption = "Speed = 0 km/h by date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Harsh Acceleration
* Values represent magnitudes of harsh acceleration. 
* Values are being shown only when speed is non-zero

```{r}

speed_valid <- subset(sensor, sensor$Speed != 0)
kable(table(speed_valid$date, speed_valid$`Harsh Acceleration*`), caption = "Harsh Acceleration values by Date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Harsh Braking
* Values represent magnitude of harsh braking
* For instance 0.31 is harsher than 0.16
* Values are being shown only when speed is non-zero

```{r, layout="l-body-outset"}
library(knitr)
library(kableExtra)

kable(table(speed_valid$date, speed_valid$`Harsh Braking*`), caption = "Harsh Braking values by Date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
##  Harsh Cornering
* Missing for all
* Values are being shown only when speed is non-zero

```{r, layout="l-body-outset"}
library(knitr)
library(kableExtra)

kable(table(speed_valid$date, speed_valid$`Harsh Cornering*`), caption = "Harsh Cornering values by Date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

## Ignition
*Dummy whether ignition was on or off
```{r, layout="l-body-outset"}
library(knitr)
library(kableExtra)

kable(table(speed_valid$date, speed_valid$`Ignition*`), caption = "Ignition (On/Off) by Date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
## Braking - Event dummy 
* Dummy whether braking occurred or not
* Not sure of this variable, does it turn on when any brake is applied or when vehicle comes to a halt?
* Need to check whether is braking dummy or harsh braking dummy
* Values are being shown for all data and for the subset of values when speed is non-zero.

```{r, layout="l-body-outset"}
library(knitr)
library(kableExtra)

kable(table(speed_valid$date, speed_valid$`Event: braking*`), caption = "Braking event by Date (Speed (Non-zero)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

kable(table(sensor$date, sensor$`Event: braking*`), caption = "Braking event by Date (Speed includes zero)") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```
## Acceleration - Event dummy 
* Dummy whether harsh acceleration happened
* Values are being shown only when speed is non-zero

```{r}
library(knitr)
library(kableExtra)

kable(table(speed_valid$date, speed_valid$`Event: acceleration*`), caption = "Speeding event by Date") %>%
  kable_styling(bootstrap_options = c("striped", "hover"))

```

# Speed over time for 19th November 2020
``````{r, fig.height=8}
subset <- subset(sensor, sensor$date == "2020-11-19")
sensor_data_19 <- ggplot(subset, aes(x=Time, y=Speed)) +
  geom_line(color="#69b3a2",  linetype=2) +
theme_ipsum() 
print(sensor_data_19)
```

# Speed tracing between 15:00 - 17:00 19th November 2020
``````{r, fig.height=8}
time_subset <- subset(subset, subset = hour %in% c(15,16))
sensor_data_15 <- ggplot(time_subset, aes(x=Time, y=Speed)) +
  geom_line(color="#69b3a2",  linetype=1) +
  ggtitle("Speed tracing 19th November 15:00-17:00") +
theme_ipsum() 
print(sensor_data_15)
```

# Map of route on 4 days

``````{r  echo=FALSE, include=FALSE}
expand <- 1.5/111.12
basemap <- get_stamenmap(bbox = c(left   = min(sensor$lon, na.rm=T) - expand, 
                                  bottom = min(sensor$lat, na.rm=T) - expand, 
                                  right  = max(sensor$lon, na.rm=T) + expand, 
                                  top    = max(sensor$lat, na.rm=T) + expand),
                         zoom = 12)

`````

``````{r, fig.height=8}
ggmap(basemap) +
  geom_point(data = sensor %>%
               filter(date %in% as.Date(c("2020-10-24",
                                          "2020-10-25",
                                          "2020-10-26",
                                          "2020-10-27"))) %>%
               arrange(Speed),
             aes(x = lon, y=lat),
             size = 1.2,
             color = "black") +
  geom_point(data = sensor %>%
               filter(date %in% as.Date(c("2020-10-24",
                                        "2020-10-25",
                                        "2020-10-26",
                                        "2020-10-27"))) %>%
               arrange(Speed),
             aes(x = lon, y=lat, color = Speed),
             size = 1) +

  labs(title = "Routes and Speeds Over Example Four Day Period\nfor Matatu KBA 970C\n",
       color = "Speed\n(km/hr)") +
  scale_color_gradientn(colours = rev(brewer.pal(n = 7, name = "Spectral"))) + 
  theme_void() +
  theme(plot.title = element_text(hjust = 0.5,
                                  face = "bold",
                                  size = 10),
        legend.position = "bottom") +
  facet_wrap(~date)
```
